Analysis of Cartesian Genetic Programming's Evolutionary Mechanisms

### 2013-06-20 CGP Journal/main.tex
Compressed the tables, fixed references

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
TODO Keywords
\end{IEEEkeywords}

@@

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
	  \hline
\textbf{Duplication} & \textbf{Ordering} & \textbf{Genome Size} & \textbf{Mutation Rate} & \textbf{MES} & \textbf{MAD} & \textbf{Active} & \textbf{Used} & \textbf{p-value} \\ \hline
\emph{Accumulate} & \emph{Normal} & 10,000 & 0.02 & 682 & 348 & 122 & 20 & 0.5602 \\ \hline
\emph{Accumulate} & \emph{Reorder}&  5,000 & 0.01 & 796 & 434 & 331 & 23 & 0.7174 \\ \hline
\rowcolor{Gray}
\emph{Skip} & \emph{Normal}       & 10,000 & 0.02 & 823 & 363 & 119 & 20 & NA \\ \hline
\emph{Skip} & \emph{Reorder}      &  2,000 & 0.01 & 870 & 362 & 198 & 20 & 0.3125 \\ \hline
\emph{Accumulate} & \emph{Dag}    &  5,000 & 0.01 & 964 & 499 & 1,174 & 22 & 0.3310 \\ \hline
\emph{Skip} & \emph{Dag}          & 10,000 & 0.02 & 1,042 & 620 & 2,083 & 24 & 0.0248 \\ \hline
\emph{Single} & \emph{Normal}     & 500 & NA & 1,214 & 497 & 39 & 17 & 0.0046 \\ \hline
\emph{Single} & \emph{Reorder}    & 200 & NA & 1,506 & 823 & 57 & 18 & 0 \\ \hline
\emph{Single} & \emph{Dag}        & 100 & NA & 2,569 & 1,313 & 43 & 15 & 0 \\ \hline
	\end{tabular}
	\caption{Parity, 5.9956882435042246e-14}
	\label{tab:parity}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
	  \hline
\textbf{Duplication} & \textbf{Ordering} & \textbf{Genome Size} & \textbf{Mutation Rate} & \textbf{MES} & \textbf{MAD} & \textbf{Active} & \textbf{Used} & \textbf{p-value} \\ \hline
\rowcolor{Gray}
\emph{Skip} & \emph{Normal}        & 10,000 & 0.005 & 16,226 & 7,455  &   283 & 49 & NA \\ \hline
\emph{Skip} & \emph{Reorder}       & 10,000 & 0.005 & 17,883 & 8,378  & 2,070 & 62 & 0.6918 \\ \hline
\emph{Accumulate} & \emph{Reorder} & 10,000 & 0.002 & 18,132 & 8,926  & 2,027 & 64 & 0.8822 \\ \hline
\emph{Accumulate} & \emph{Normal}  & 10,000 & 0.005 & 21,428 & 6,379  &   279 & 49 & 0.2525 \\ \hline
\emph{Accumulate} & \emph{Dag}     & 10,000 & 0.002 & 24,855 & 12,512 & 4,059 & 69 & 0.0736 \\ \hline
\emph{Skip} & \emph{Dag}           & 10,000 & 0.002 & 25,678 & 12,034 & 4,178 & 62 & 0.0211 \\ \hline
\emph{Single} & \emph{Reorder}     & 100    & NA    & 26,558 & 7,858  &    52 & 32 & 0.0076 \\ \hline
\emph{Single} & \emph{Normal}      & 2,000  & NA    & 31,057 & 14,475 &   141 & 43 & 0.0085 \\ \hline
\emph{Single} & \emph{Dag}         & 100    & NA    & 34,208 & 10,069 &    59 & 34 & 0 \\ \hline
	\end{tabular}
	\caption{Encode 1.1926088284917749e-06}
	\label{tab:encode}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
	  \hline
\textbf{Duplication} & \textbf{Ordering} & \textbf{Genome Size} & \textbf{Mutation Rate} & \textbf{MES} & \textbf{MAD} & \textbf{Active} & \textbf{Used} & \textbf{p-value} \\ \hline
\emph{Skip} & \emph{Reorder}       &  2,000 & 0.002 &  62,853 & 20,023 &   757 & 114 & 0.5247 \\ \hline
\emph{Single} & \emph{Reorder}     &    500 &    NA &  63,731 & 13,119 &   245 &  97 & 0.3610 \\ \hline
\emph{Accumulate} & \emph{Normal}  &  1,000 & 0.002 &  67,390 & 11,991 &   211 &  87 & 0.8767 \\ \hline
\rowcolor{Gray}
\emph{Skip} & \emph{Normal}        &  1,000 & 0.002 &  68,231 & 15,115 &   202 &  85 & NA \\ \hline
\emph{Single} & \emph{Normal}      &  1,000 &    NA &  68,819 & 14,590 &   203 &  85 & 0.7695 \\ \hline
\emph{Accumulate} & \emph{Reorder} &  2,000 & 0.002 &  72,991 & 18,650 &   760 & 115 & 0.1868 \\ \hline
\emph{Skip} & \emph{Dag}           & 10,000 & 0.001 & 132,821 & 29,432 & 4,983 & 125 & 0 \\ \hline
\emph{Accumulate} & \emph{Dag}     &  5,000 & 0.002 & 133,188 & 29,326 & 2,542 & 125 & 0 \\ \hline
\emph{Single} & \emph{Dag}         &    500 &    NA & 160,696 & 39,242 &   299 &  97 & 0 \\ \hline
	\end{tabular}
	\caption{Decode 1.0182688124966139e-36}
	\label{tab:decode}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
	  \hline
\textbf{Duplication} & \textbf{Ordering} & \textbf{Genome Size} & \textbf{Mutation Rate} & \textbf{MES} & \textbf{MAD} & \textbf{Active} & \textbf{Used} & \textbf{p-value} \\ \hline
\emph{Skip} & \emph{Reorder}      & 10,000 & 0.0005  & 196,116  & 60,326  & 3,175 & 484 & 0  \\ \hline
\emph{Accumulate} & \emph{Reorder}& 10,000 & 0.001 & 199,193 & 85,970 & 3,117 & 489 & 0  \\ \hline
\emph{Single} & \emph{Dag} & 200  & NA & 239,844 & 78,739 & 114 & 74 & 0 \\ \hline
\emph{Single} & \emph{Reorder}    & 500   & NA & 243,664 & 91,935 & 266 & 131 & 0 \\ \hline
\emph{Skip} & \emph{Dag}          & 2,000 & 0.002 & 369,931 & 174,408 & 952 & 149 & 0.0100 \\ \hline
\emph{Accumulate} & \emph{Dag}    & 2,000 & 0.002  & 379,202 & 147,289 & 973 & 158 & 0.0053  \\ \hline
\emph{Single} & \emph{Normal}     & 5,000  & NA & 399,911 & 194,384 & 273 & 112 & 0.0047 \\ \hline
\emph{Accumulate} & \emph{Normal} & 2,000 & 0.002 & 483,938 & 227,686 & 168 & 87 & 0.45 \\ \hline
\rowcolor{Gray}
\emph{Skip} & \emph{Normal} & 2,000 & 0.002 & 657,121 & 311,376 & 175 & 84 & NA \\ \hline

	\end{tabular}
	\caption{Multiply 5.1429910779824196e-18}
	\label{tab:multiply}
\end{table*}

The results from running each tuned parameter configuration for each algorithm
combination on each problem are summarized by Table~\ref{tab:parity},
Table~\ref{tab:encode}, Table~\ref{tab:decode}, and Table~\ref{tab:multiply}.
Rows in each table are ordered by that algorithm combination's median evaluations to
success (MES), with the best algorithm combination appearing at the top of each table.
The highlighted row marks \emph{Skip Normal}, the control algorithm combination
representing traditional CGP's performance.  All configurations for each problem
were then testing using Kruskal-Wallis to determine if any configuration is
statistically different from any other.  All problems received highly significant
values.  As such, each algorithm combination was then compared with the control
to determine statistical significance using pairwise Mann-Whitney~U tests. The
resulting p-value for each test is reported in the table.

@@

By ordering the algorithm combinations by MES and examining that ordering across each
problem, we can look for general trends.  From the tables, it appears on easier problems
which method of duplication is in use is the strongest predictor of rank, while
the ordering methods are more important on harder problems.  This comes from the
observation that in for Parity and Encode (Table~\ref{tab:parity} and Table~\ref{tab:encode})
the first column is effectively sorted, while for Decode and Multiply
(Table~\ref{tab:decode} and Table~\ref{tab:multiply}) the second column is sorted.
Each table includes coincidentally one exception to these rules.  On Parity and
Encode pairings with \emph{DAG} are ranked lower than other ordering methods, so much
so as to break the ordering once on both problems.  On Decode \emph{Reorder Duplicate}
ranks worse than expected.  Finally on Multiply \emph{Single DAG} does far better
than expected.

@@

Using the algorithm given in Figure~\ref{fig:simplify}, we can determine the
absolute minimum set of nodes in a genome's solution that are necessary to
reproduce the original output of that genome.  This allows us to discuss
the usefulness of each node as either being part of the minimum solution, a
duplicate of a node in the minimum solution, or irrelevant to the solution.
As such we can then examine how redundancy is being used, how frequently useful
structures exist in the inactive nodes, and how useful nodes are constructed by evolution.
This algorithm can also be used to construct simplified genomes when combined
with the algorithm given in Figure~\ref{fig:reorder}.
\end{comment}
\subsection{Never Active Nodes}
%TODO Somewhere in here you should talk about how Normal has the least number of active,
%which is probably related to how many never become active
In line with the suggestions made in~\cite{miller:2006:redundancy}, the evolved
number of active nodes across all problems and all algorithm configurations
in Table~\ref{tab:parity}, Table~\ref{tab:encode}, Table~\ref{tab:decode}, and Table~\ref{tab:multiply}
was significantly lower than the genome size.  With the exception of a few combinations
with \emph{DAG} and \emph{Single}, the evolved results were over 50\% inactive.
\emph{Skip Normal}, our control representing normal CGP, had only 1\%, 3\%, 20\%,
and 9\% of nodes in the genome active in the median runs of Parity, Encode, Decode,
and Multiply, respectively.  These final results fall in line with the predictions
made in~\cite{goldman:2013:ordering}, in that if the problems are ordered based on
their number of output locations, their evolved percentage of active nodes is also ordered.

@@

In CGP it is common to report what fraction of the genome is active, as this represents
the easiest estimate of solution size.  By utilizing \emph{footprint} comparison
and the \Call{Simplify}{} algorithm, we can get a much better estimate of what
purpose each node in the genome serves.
Table~\ref{tab:parity}, Table~\ref{tab:encode}, Table~\ref{tab:decode},
and Table~\ref{tab:multiply} provide two column relevant to node behavior: Active
and Used.  The former gives the traditional measure of the median amount of
active nodes in the solution found.  The latter is the median number of nodes
still active after applying the \Call{Simplify}{} algorithm.

@@

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	  \hline
\textbf{Duplicate} & \textbf{Ordering} & \textbf{Active Unchanged} & \textbf{No Reactivation} & \textbf{Reactivated Changed} \\ \hline
       \emph{Skip} &     \emph{DAG} & 0.00\% & 3.48\% & 55.63\% \\ \hline
 \emph{Accumulate} &     \emph{DAG} & 0.00\% & 8.48\% & 52.82\% \\ \hline
 \emph{Accumulate} & \emph{Reorder} & 2.09\% & 33.26\% & 58.46\% \\ \hline
   \rowcolor{Gray}
       \emph{Skip} &  \emph{Normal} & 2.25\% & 39.35\% & 67.17\% \\ \hline
 \emph{Accumulate} &  \emph{Normal} & 2.69\% & 39.99\% & 67.27\% \\ \hline
       \emph{Skip} & \emph{Reorder} & 11.46\% & 53.99\% & 59.47\% \\ \hline
     \emph{Single} & \emph{Reorder} & 24.94\% & 80.49\% & 60.63\% \\ \hline
     \emph{Single} &  \emph{Normal} & 26.44\% & 83.55\% & 59.96\% \\ \hline
     \emph{Single} &     \emph{DAG} & 27.80\% & 83.73\% & 59.92\% \\ \hline
	\end{tabular}
	\caption{Parity}
	\label{tab:parity_reactivate}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	  \hline
\textbf{Duplicate} & \textbf{Ordering} & \textbf{Active Unchanged} & \textbf{No Reactivation} & \textbf{Reactivated Changed} \\ \hline
 \emph{Accumulate} &     \emph{DAG} & 0.05\% & 4.06\% & 48.70\% \\ \hline
       \emph{Skip} &     \emph{DAG} & 0.06\% & 3.85\% & 48.33\% \\ \hline
 \emph{Accumulate} & \emph{Reorder} & 1.77\% & 19.24\% & 55.47\% \\ \hline
       \emph{Skip} & \emph{Reorder} & 2.02\% & 19.80\% & 56.15\% \\ \hline
   \rowcolor{Gray}
       \emph{Skip} &  \emph{Normal} & 14.16\% & 54.78\% & 64.14\% \\ \hline
 \emph{Accumulate} &  \emph{Normal} & 14.52\% & 54.79\% & 64.17\% \\ \hline
     \emph{Single} &  \emph{Normal} & 33.77\% & 79.34\% & 59.86\% \\ \hline
     \emph{Single} &     \emph{DAG} & 35.98\% & 80.91\% & 58.39\% \\ \hline
     \emph{Single} & \emph{Reorder} & 39.95\% & 78.91\% & 59.72\% \\ \hline
	\end{tabular}
	\caption{Encode}
	\label{tab:encode_reactivate}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	  \hline
\textbf{Duplicate} & \textbf{Ordering} & \textbf{Active Unchanged} & \textbf{No Reactivation} & \textbf{Reactivated Changed} \\ \hline
 \emph{Accumulate} &     \emph{DAG} & 0.03\% & 10.10\% & 54.99\% \\ \hline
       \emph{Skip} &     \emph{DAG} & 0.05\% & 10.21\% & 51.33\% \\ \hline
 \emph{Accumulate} & \emph{Reorder} & 9.98\% & 49.65\% & 70.57\% \\ \hline
       \emph{Skip} & \emph{Reorder} & 12.53\% & 51.63\% & 70.24\% \\ \hline
     \emph{Single} & \emph{Reorder} & 28.38\% & 79.73\% & 66.85\% \\ \hline
     \emph{Single} &  \emph{Normal} & 29.43\% & 80.86\% & 65.85\% \\ \hline
     \emph{Single} &     \emph{DAG} & 32.60\% & 81.73\% & 57.70\% \\ \hline
 \emph{Accumulate} &  \emph{Normal} & 51.11\% & 84.54\% & 72.57\% \\ \hline
   \rowcolor{Gray}
       \emph{Skip} &  \emph{Normal} & 65.47\% & 89.17\% & 72.21\% \\ \hline

	\end{tabular}
	\caption{Decode}
	\label{tab:decode_reactivate}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	  \hline
\textbf{Duplicate} & \textbf{Ordering} & \textbf{Active Unchanged} & \textbf{No Reactivation} & \textbf{Reactivated Changed} \\ \hline
 \emph{Accumulate} & \emph{Reorder} & 1.14\% & 22.09\% & 71.75\% \\ \hline
 \emph{Accumulate} &     \emph{DAG} & 4.52\% & 45.38\% & 63.42\% \\ \hline
       \emph{Skip} &     \emph{DAG} & 5.05\% & 45.58\% & 63.03\% \\ \hline
       \emph{Skip} & \emph{Reorder} & 10.40\% & 47.67\% & 71.79\% \\ \hline
     \emph{Single} &  \emph{Normal} & 18.85\% & 78.19\% & 70.11\% \\ \hline
     \emph{Single} &     \emph{DAG} & 20.72\% & 80.20\% & 65.10\% \\ \hline
     \emph{Single} & \emph{Reorder} & 22.14\% & 80.21\% & 69.76\% \\ \hline
 \emph{Accumulate} &  \emph{Normal} & 57.69\% & 87.27\% & 78.95\% \\ \hline
   \rowcolor{Gray}
       \emph{Skip} &  \emph{Normal} & 73.46\% & 92.04\% & 79.07\% \\ \hline
	\end{tabular}
	\caption{Multiply}
	\label{tab:multiply_reactivate}
\end{table*}

@@

Table~\ref{tab:parity_reactivate}, Table~\ref{tab:encode_reactivate},
Table~\ref{tab:decode_reactivate}, and Table~\ref{tab:multiply_reactivate} provide a
novel look at CGP's replacement behavior.  Each time an offspring replaced a parent
in the final runs, we recorded how the behavior of each node differed between the
parent and the offspring.

@@

\appendices
\section{Example Appendix}
Appendix one text goes here.

@@

% use section* for acknowledgement
\section*{Acknowledgment}

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
TODO The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Cartesian genetic programming (CGP), Analysis
\end{IEEEkeywords}

@@

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
	  \hline
	  \multicolumn{9}{|c|}{\textbf{3-bit Parity}, Kruskal-Wallis=$5.99\cdot 10^{-14}$} \\ \hline
\textbf{Duplication} & \textbf{Ordering} & \textbf{Genome Size} & \textbf{Mutation Rate} & \textbf{MES} & \textbf{MAD} & \textbf{Active} & \textbf{Used} & \textbf{p-value} \\ \hline
\emph{Accumulate} & \emph{Normal} & 10,000 & 0.02 & 682 & 348 & 122 & 20 & 0.5602 \\ \hline
\emph{Accumulate} & \emph{Reorder}&  5,000 & 0.01 & 796 & 434 & 331 & 23 & 0.7174 \\ \hline
\rowcolor{Gray}
\emph{Skip} & \emph{Normal}       & 10,000 & 0.02 & 823 & 363 & 119 & 20 & NA \\ \hline
\emph{Skip} & \emph{Reorder}      &  2,000 & 0.01 & 870 & 362 & 198 & 20 & 0.3125 \\ \hline
\emph{Accumulate} & \emph{Dag}    &  5,000 & 0.01 & 964 & 499 & 1,174 & 22 & 0.3310 \\ \hline
\emph{Skip} & \emph{Dag}          & 10,000 & 0.02 & 1,042 & 620 & 2,083 & 24 & 0.0248 \\ \hline
\emph{Single} & \emph{Normal}     & 500 & NA & 1,214 & 497 & 39 & 17 & 0.0046 \\ \hline
\emph{Single} & \emph{Reorder}    & 200 & NA & 1,506 & 823 & 57 & 18 & 0 \\ \hline
\emph{Single} & \emph{Dag}        & 100 & NA & 2,569 & 1,313 & 43 & 15 & 0 \\ \hline\hline

	  \multicolumn{9}{|c|}{\textbf{16-4-bit Encode}, Kruskal-Wallis=$1.19\cdot 10^{-6}$} \\ \hline

\textbf{Duplication} & \textbf{Ordering} & \textbf{Genome Size} & \textbf{Mutation Rate} & \textbf{MES} & \textbf{MAD} & \textbf{Active} & \textbf{Used} & \textbf{p-value} \\ \hline
\rowcolor{Gray}
\emph{Skip} & \emph{Normal}        & 10,000 & 0.005 & 16,226 & 7,455  &   283 & 49 & NA \\ \hline
\emph{Skip} & \emph{Reorder}       & 10,000 & 0.005 & 17,883 & 8,378  & 2,070 & 62 & 0.6918 \\ \hline
\emph{Accumulate} & \emph{Reorder} & 10,000 & 0.002 & 18,132 & 8,926  & 2,027 & 64 & 0.8822 \\ \hline
\emph{Accumulate} & \emph{Normal}  & 10,000 & 0.005 & 21,428 & 6,379  &   279 & 49 & 0.2525 \\ \hline
\emph{Accumulate} & \emph{Dag}     & 10,000 & 0.002 & 24,855 & 12,512 & 4,059 & 69 & 0.0736 \\ \hline
\emph{Skip} & \emph{Dag}           & 10,000 & 0.002 & 25,678 & 12,034 & 4,178 & 62 & 0.0211 \\ \hline
\emph{Single} & \emph{Reorder}     & 100    & NA    & 26,558 & 7,858  &    52 & 32 & 0.0076 \\ \hline
\emph{Single} & \emph{Normal}      & 2,000  & NA    & 31,057 & 14,475 &   141 & 43 & 0.0085 \\ \hline
\emph{Single} & \emph{Dag}         & 100    & NA    & 34,208 & 10,069 &    59 & 34 & 0 \\ \hline\hline

	  \multicolumn{9}{|c|}{\textbf{4-16-bit Decode}, Kruskal-Wallis=$1.02\cdot 10^{-36}$} \\ \hline

\textbf{Duplication} & \textbf{Ordering} & \textbf{Genome Size} & \textbf{Mutation Rate} & \textbf{MES} & \textbf{MAD} & \textbf{Active} & \textbf{Used} & \textbf{p-value} \\ \hline
\emph{Skip} & \emph{Reorder}       &  2,000 & 0.002 &  62,853 & 20,023 &   757 & 114 & 0.5247 \\ \hline
\emph{Single} & \emph{Reorder}     &    500 &    NA &  63,731 & 13,119 &   245 &  97 & 0.3610 \\ \hline
\emph{Accumulate} & \emph{Normal}  &  1,000 & 0.002 &  67,390 & 11,991 &   211 &  87 & 0.8767 \\ \hline
\rowcolor{Gray}
\emph{Skip} & \emph{Normal}        &  1,000 & 0.002 &  68,231 & 15,115 &   202 &  85 & NA \\ \hline
\emph{Single} & \emph{Normal}      &  1,000 &    NA &  68,819 & 14,590 &   203 &  85 & 0.7695 \\ \hline
\emph{Accumulate} & \emph{Reorder} &  2,000 & 0.002 &  72,991 & 18,650 &   760 & 115 & 0.1868 \\ \hline
\emph{Skip} & \emph{Dag}           & 10,000 & 0.001 & 132,821 & 29,432 & 4,983 & 125 & 0 \\ \hline
\emph{Accumulate} & \emph{Dag}     &  5,000 & 0.002 & 133,188 & 29,326 & 2,542 & 125 & 0 \\ \hline
\emph{Single} & \emph{Dag}         &    500 &    NA & 160,696 & 39,242 &   299 &  97 & 0 \\ \hline\hline

	  \multicolumn{9}{|c|}{\textbf{3-bit Multiply}, Kruskal-Wallis=$5.14\cdot 10^{-18}$} \\ \hline

\textbf{Duplication} & \textbf{Ordering} & \textbf{Genome Size} & \textbf{Mutation Rate} & \textbf{MES} & \textbf{MAD} & \textbf{Active} & \textbf{Used} & \textbf{p-value} \\ \hline
\emph{Skip} & \emph{Reorder}      & 10,000 & 0.0005  & 196,116  & 60,326  & 3,175 & 484 & 0  \\ \hline
\emph{Accumulate} & \emph{Reorder}& 10,000 & 0.001 & 199,193 & 85,970 & 3,117 & 489 & 0  \\ \hline
\emph{Single} & \emph{Dag} & 200  & NA & 239,844 & 78,739 & 114 & 74 & 0 \\ \hline
\emph{Single} & \emph{Reorder}    & 500   & NA & 243,664 & 91,935 & 266 & 131 & 0 \\ \hline
\emph{Skip} & \emph{Dag}          & 2,000 & 0.002 & 369,931 & 174,408 & 952 & 149 & 0.0100 \\ \hline
\emph{Accumulate} & \emph{Dag}    & 2,000 & 0.002  & 379,202 & 147,289 & 973 & 158 & 0.0053  \\ \hline
\emph{Single} & \emph{Normal}     & 5,000  & NA & 399,911 & 194,384 & 273 & 112 & 0.0047 \\ \hline
\emph{Accumulate} & \emph{Normal} & 2,000 & 0.002 & 483,938 & 227,686 & 168 & 87 & 0.45 \\ \hline
\rowcolor{Gray}
\emph{Skip} & \emph{Normal} & 2,000 & 0.002 & 657,121 & 311,376 & 175 & 84 & NA \\ \hline

	\end{tabular}
	\caption{Results for algorithm configuration on each problem.  Highlighted row
	is the control configuration \emph{Skip Normal}.  MES=Median Evaluations to Success; MAD=Median
	Absolute Deviation; p-value=Result from Mann-Whitney~U comparison with control.
	}
	\label{tab:results}
\end{table*}

The results from running each tuned parameter configuration for each algorithm
combination on each problem are summarized by Table~\ref{tab:results}.
Rows in each table are ordered by that algorithm combination's median evaluations to
success (MES), with the best algorithm combination appearing at the top of each table.
The highlighted row marks \emph{Skip Normal}, the control algorithm combination
representing traditional CGP's performance.  All configurations for each problem
were then testing using Kruskal-Wallis to determine if any configuration is
statistically different from any other.  All problems received highly significant
values.  As such, each algorithm combination was then compared with the control
to determine statistical significance using pairwise Mann-Whitney~U tests. The
resulting p-value for each test is reported in the table.

@@

By ordering the algorithm combinations by MES and examining that ordering across each
problem, we can look for general trends.  From the tables, it appears on easier problems
which method of duplication is in use is the strongest predictor of rank, while
the ordering methods are more important on harder problems.  This comes from the
observation that in for Parity and Encode (Upper half of Table~\ref{tab:results})
the first column is effectively sorted, while for Decode and Multiply
(Lower half of Table~\ref{tab:results})
the second column is sorted.
Each table includes coincidentally one exception to these rules.  On Parity and
Encode pairings with \emph{DAG} are ranked lower than other ordering methods, so much
so as to break the ordering once on both problems.  On Decode \emph{Reorder Duplicate}
ranks worse than expected.  Finally on Multiply \emph{Single DAG} does far better
than expected.

@@

Using the algorithm given in Figure~\ref{fig:simplify}, we can determine the
absolute minimum set of nodes in a genome's solution that are necessary to
reproduce the original output of that genome.  This allows us to discuss
the usefulness of each node as either being part of the minimum solution, a
duplicate of a node in the minimum solution, or irrelevant to the solution.
As such we can then examine how redundancy is being used, how frequently useful
structures exist in the inactive nodes, and how useful nodes are constructed by evolution.
This algorithm can also be used to construct simplified genomes when combined
with the algorithm given in Figure~\ref{fig:reorder}.
\end{comment}
\subsection{Never Active Nodes}
%TODO Somewhere in here you should talk about how Normal has the least number of active,
%which is probably related to how many never become active
In line with the suggestions made in~\cite{miller:2006:redundancy}, the evolved
number of active nodes across all problems and all algorithm configurations
in Table~\ref{tab:results}
was significantly lower than the genome size.  With the exception of a few combinations
with \emph{DAG} and \emph{Single}, the evolved results were over 50\% inactive.
\emph{Skip Normal}, our control representing normal CGP, had only 1\%, 3\%, 20\%,
and 9\% of nodes in the genome active in the median runs of Parity, Encode, Decode,
and Multiply, respectively.  These final results fall in line with the predictions
made in~\cite{goldman:2013:ordering}, in that if the problems are ordered based on
their number of output locations, their evolved percentage of active nodes is also ordered.

@@

In CGP it is common to report what fraction of the genome is active, as this represents
the easiest estimate of solution size.  By utilizing \emph{footprint} comparison
and the \Call{Simplify}{} algorithm, we can get a much better estimate of what
purpose each node in the genome serves.
Table~\ref{tab:results} provides two columns relevant to node behavior: Active
and Used.  The former gives the traditional measure of the median amount of
active nodes in the solution found.  The latter is the median number of nodes
still active after applying the \Call{Simplify}{} algorithm.

@@

\begin{table*}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	  \hline
	  \multicolumn{5}{|c|}{\textbf{3-bit Parity}} \\ \hline
\textbf{Duplicate} & \textbf{Ordering} & \textbf{Active Unchanged} & \textbf{No Reactivation} & \textbf{Reactivated Changed} \\ \hline
       \emph{Skip} &     \emph{DAG} & 0.00\% & 3.48\% & 55.63\% \\ \hline
 \emph{Accumulate} &     \emph{DAG} & 0.00\% & 8.48\% & 52.82\% \\ \hline
 \emph{Accumulate} & \emph{Reorder} & 2.09\% & 33.26\% & 58.46\% \\ \hline
   \rowcolor{Gray}
       \emph{Skip} &  \emph{Normal} & 2.25\% & 39.35\% & 67.17\% \\ \hline
 \emph{Accumulate} &  \emph{Normal} & 2.69\% & 39.99\% & 67.27\% \\ \hline
       \emph{Skip} & \emph{Reorder} & 11.46\% & 53.99\% & 59.47\% \\ \hline
     \emph{Single} & \emph{Reorder} & 24.94\% & 80.49\% & 60.63\% \\ \hline
     \emph{Single} &  \emph{Normal} & 26.44\% & 83.55\% & 59.96\% \\ \hline
     \emph{Single} &     \emph{DAG} & 27.80\% & 83.73\% & 59.92\% \\ \hline\hline

  \multicolumn{5}{|c|}{\textbf{16-4-bit Encode}} \\ \hline

\textbf{Duplicate} & \textbf{Ordering} & \textbf{Active Unchanged} & \textbf{No Reactivation} & \textbf{Reactivated Changed} \\ \hline
 \emph{Accumulate} &     \emph{DAG} & 0.05\% & 4.06\% & 48.70\% \\ \hline
       \emph{Skip} &     \emph{DAG} & 0.06\% & 3.85\% & 48.33\% \\ \hline
 \emph{Accumulate} & \emph{Reorder} & 1.77\% & 19.24\% & 55.47\% \\ \hline
       \emph{Skip} & \emph{Reorder} & 2.02\% & 19.80\% & 56.15\% \\ \hline
   \rowcolor{Gray}
       \emph{Skip} &  \emph{Normal} & 14.16\% & 54.78\% & 64.14\% \\ \hline
 \emph{Accumulate} &  \emph{Normal} & 14.52\% & 54.79\% & 64.17\% \\ \hline
     \emph{Single} &  \emph{Normal} & 33.77\% & 79.34\% & 59.86\% \\ \hline
     \emph{Single} &     \emph{DAG} & 35.98\% & 80.91\% & 58.39\% \\ \hline
     \emph{Single} & \emph{Reorder} & 39.95\% & 78.91\% & 59.72\% \\ \hline\hline

  \multicolumn{5}{|c|}{\textbf{4-16-bit Decode}} \\ \hline

\textbf{Duplicate} & \textbf{Ordering} & \textbf{Active Unchanged} & \textbf{No Reactivation} & \textbf{Reactivated Changed} \\ \hline
 \emph{Accumulate} &     \emph{DAG} & 0.03\% & 10.10\% & 54.99\% \\ \hline
       \emph{Skip} &     \emph{DAG} & 0.05\% & 10.21\% & 51.33\% \\ \hline
 \emph{Accumulate} & \emph{Reorder} & 9.98\% & 49.65\% & 70.57\% \\ \hline
       \emph{Skip} & \emph{Reorder} & 12.53\% & 51.63\% & 70.24\% \\ \hline
     \emph{Single} & \emph{Reorder} & 28.38\% & 79.73\% & 66.85\% \\ \hline
     \emph{Single} &  \emph{Normal} & 29.43\% & 80.86\% & 65.85\% \\ \hline
     \emph{Single} &     \emph{DAG} & 32.60\% & 81.73\% & 57.70\% \\ \hline
 \emph{Accumulate} &  \emph{Normal} & 51.11\% & 84.54\% & 72.57\% \\ \hline
   \rowcolor{Gray}
       \emph{Skip} &  \emph{Normal} & 65.47\% & 89.17\% & 72.21\% \\ \hline\hline

  \multicolumn{5}{|c|}{\textbf{3-bit Multiply}} \\ \hline

\textbf{Duplicate} & \textbf{Ordering} & \textbf{Active Unchanged} & \textbf{No Reactivation} & \textbf{Reactivated Changed} \\ \hline
 \emph{Accumulate} & \emph{Reorder} & 1.14\% & 22.09\% & 71.75\% \\ \hline
 \emph{Accumulate} &     \emph{DAG} & 4.52\% & 45.38\% & 63.42\% \\ \hline
       \emph{Skip} &     \emph{DAG} & 5.05\% & 45.58\% & 63.03\% \\ \hline
       \emph{Skip} & \emph{Reorder} & 10.40\% & 47.67\% & 71.79\% \\ \hline
     \emph{Single} &  \emph{Normal} & 18.85\% & 78.19\% & 70.11\% \\ \hline
     \emph{Single} &     \emph{DAG} & 20.72\% & 80.20\% & 65.10\% \\ \hline
     \emph{Single} & \emph{Reorder} & 22.14\% & 80.21\% & 69.76\% \\ \hline
 \emph{Accumulate} &  \emph{Normal} & 57.69\% & 87.27\% & 78.95\% \\ \hline
   \rowcolor{Gray}
       \emph{Skip} &  \emph{Normal} & 73.46\% & 92.04\% & 79.07\% \\ \hline
	\end{tabular}
	\caption{Behavior of offspring replacement for each algorithm combination
	         on each problem.}
	\label{tab:reactivate}
\end{table*}

@@

Table~\ref{tab:reactivate} provides a
novel look at CGP's replacement behavior.  Each time an offspring replaced a parent
in the final runs, we recorded how the behavior of each node differed between the
parent and the offspring.

@@

\begin{comment}
\appendices
\section{Example Appendix}
Appendix one text goes here.

@@

\end{comment}
% use section* for acknowledgement
\section*{Acknowledgment}
